---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---


## Loading and preprocessing the data


```{r load_data}
unzip("activity.zip")
data_activity <- read.csv("activity.csv")

```


## What is mean total number of steps taken per day?

### Histogram of the total number of steps taken each day

```{r plot_hist_steps_day}

nb_step_by_days <- tapply(data_activity$steps,data_activity$date,sum)
hist(nb_step_by_days, main = "Total number of steps each day")
```

### Mean and median number of steps taken each day


```{r mean_median}
options(scipen = 2)
s_step_day <- summary(nb_step_by_days)
mean_step_day <- s_step_day["Mean"]
median_step_day <- s_step_day["Median"]

```
The mean of the number of steps taken each day is `r mean_step_day` and the median is `r median_step_day`.


## What is the average daily activity pattern?

### Time series plot of the average number of steps taken



```{r plot_daily_pattern}

steps_by_interval <- with(data_activity, tapply(steps,interval,function(x) sum(x,na.rm = TRUE)))
plot(names(steps_by_interval), steps_by_interval, main = "Daily average pattern", type ="l", ylab = "Number of steps", xlab = "Interval identifier")
```

### The 5-minute interval that, on average, contains the maximum number of steps

The 5-minute interval that, on average, contains the maximum number of steps is `r names(which.max(steps_by_interval))`.

## Imputing missing values

### Strategy for imputing missing data

There are `r sum(is.na(data_activity$steps))` missing values.

These missing values are replaced with the average value over the 5 min interval.

```{r replace_missing_values}

mean_steps_by_interval <- tapply(data_activity$steps,data_activity$interval,function(x) mean(x,na.rm = TRUE))

steps_missing <- is.na(data_activity$steps)
data_activity[steps_missing, "steps"] <- mean_steps_by_interval[as.character(data_activity[steps_missing, "interval"])]

```

### Histogram of the total number of steps taken each day after missing values are imputed.


```{r hist_steps_day_naremoved}
nb_step_by_days_narm <- tapply(data_activity$steps,data_activity$date,sum)
hist(nb_step_by_days_narm, main = "Total number of steps each day", xlab = "Number of steps per day")

```

### Mean and medain number of steps taken each day once N.A. are replaced.

```{r mean_median_na}
options(scipen = 2)
s_step_day_narm <- summary(nb_step_by_days_narm)
mean_step_day_narm <- s_step_day_narm["Mean"]
median_step_day_narm <- s_step_day_narm["Median"]

```
The mean of the number of steps taken each day is `r mean_step_day_narm` and the median is `r median_step_day_narm`. Only the value for the median slightly differs with the value computed with NA(mean: `r mean_step_day`, median: `r median_step_day`).


## Are there differences in activity patterns between weekdays and weekends?


Create factor with two levels weekday and weekend.

```{r create_weekday_factor}

wdays <- strptime(data_activity$date, format ="%Y-%m-%d")$wday %in% c(0,6)
wdays <- as.factor(wdays)
levels(wdays) <- c("weekday", "weekend")

data_activity <- cbind(data_activity, "days" = wdays)

```

Create panel plot.

```{r create_panel_plot}
library(reshape2)
steps_by_interval_w <- with(data_activity, tapply(steps,list(days,interval),sum))
steps_by_interval_wl <- melt(steps_by_interval_w, varnames = c("day","interval"))

library(lattice)
xyplot(value ~ interval|day, steps_by_interval_wl, layout = c(1,2), type = "l", xlab = "Interval", ylab = "Number of steps")

```
